Detecting Stylistic Variation in Pop Production using the Spotify API

### 2023-11-09 outputs/paper/paper.qmd
artists folder, wrestling with latex

library(tidyverse)
library(lubridate)

@@

We find that danceability is the strongest predictor of whether or not a song has been produced by Antonoff. Furthermore, when the variance is considered along with other features of the dataset, the variance tends to be strongest classifier of whether a song was produced by Antonoff. Daanceability in concert with other features in the dataset provides the best predictions.

@@

In total 3738 tracks from X different artists. (Producers, etc)

@@

```{r}
#| label: fig-quadvalence
#| fig-cap: Valence and Energy Quadrant
#| echo: false
#| warning: false

# df <- read_csv(here::here("inputs/data/clean_df.csv"))
# 
# ggplot(df, aes(x = energy, y = valence)) +
#   geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
#   geom_vline(xintercept = mean(df$energy), color = "gray") +
#   geom_hline(yintercept = mean(df$valence), color = "gray") +
#   theme_minimal()
```

@@

```{r}
#| label: fig-quaddanceability2
#| fig-cap: Danceability and Energy Quadrant
#| echo: false
#| warning: false
# 
# ggplot(df, aes(x = energy, y = danceability)) +
#   geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
#   geom_vline(xintercept = mean(df$energy), color = "gray") +
#   geom_hline(yintercept = mean(df$danceability),
#              color = "gray") +
#   theme_minimal()
```

@@

# data_reduced <- read_csv(here::here("inputs/data/data_reduced.csv"))
# 
# data_reduced |>
#   pivot_longer(
#     cols = c(
#       danceability,
#       energy,
#       loudness,
#       mode,
#       speechiness,
#       acousticness,
#       instrumentalness,
#       liveness,
#       valence,
#       tempo
#     ),
#     names_to = "name",
#     values_to = "value"
#   ) |>
#   ggplot(aes(x = value, color = jack)) +
#   stat_ecdf() +
#   facet_wrap(vars(name),
#              nrow = 3,
#              ncol = 4,
#              scales = "free") +
#   theme_minimal() +
#   labs(x = "Value",
#        y = "Proportion")

#even tho there's similrities, there's differences
```

@@

# df <- read_csv(here::here("inputs/data/df.csv"))
# 
# df <- df |> 
#   mutate(jack = factor(jack))
# 
# df |>
#   filter(album_release_date_precision == "day") |>
#   mutate(album_release_date = ymd(album_release_date)) |>
#   ggplot(aes(y = energy, x = album_release_date, color = jack)) +
#   geom_point(alpha = 0.3) +
#   geom_smooth(method = lm, se = FALSE) +
#   theme_minimal() +
#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Energy") +
#   facet_wrap(~ artist_name, ncol = 3)
```

@@

# df |>
#   filter(album_release_date_precision == "day") |>
#   mutate(album_release_date = ymd(album_release_date)) |>
#   ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
#   geom_point(alpha = 0.3) +
#   geom_smooth(method = lm, se = FALSE) +
#   theme_minimal() +
#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Danceability") +
#   facet_wrap( ~ artist_name, ncol = 3)
```

@@

# df |>
#   filter(album_release_date_precision == "day") |>
#   mutate(album_release_date = ymd(album_release_date)) |>
#   ggplot(aes(y = valence, x = album_release_date, color = jack)) +
#   geom_point(alpha = 0.3) +
#   geom_smooth(method = lm, se = FALSE) +
#   theme_minimal() +
#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Valence") +
#   facet_wrap( ~ artist_name, ncol = 3)
```

@@

# df |>
#   filter(album_release_date_precision == "day") |>
#   mutate(album_release_date = ymd(album_release_date)) |>
#   ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
#   geom_point(alpha = 0.3) +
#   geom_smooth(method = lm, se = FALSE) +
#   theme_minimal() +
#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Acousticness") +
#   facet_wrap( ~ artist_name, ncol = 3)
```

@@

# df |>
#   filter(album_release_date_precision == "day") |>
#   mutate(album_release_date = ymd(album_release_date)) |>
#   ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
#   geom_point(alpha = 0.3) +
#   geom_smooth(method = lm, se = FALSE) +
#   theme_minimal() +
#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Instrumentalness") +
#   facet_wrap( ~ artist_name, ncol = 3)
```

@@

# df |>
#   filter(album_release_date_precision == "day") |>
#   mutate(album_release_date = ymd(album_release_date)) |>
#   ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
#   geom_point(alpha = 0.3) +
#   geom_smooth(method = lm, se = FALSE) +
#   theme_minimal() +
#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Tempo") +
#   facet_wrap( ~ artist_name, ncol = 3)
```

@@

$$
\mbox{Pr}(y_i=1) = \mbox{logit}^{-1}\left(\beta_0+\beta_1 x_i\right)
$$

where $\text{Pr}(y_i=1)$ is the probability of a song being produced by Antonoff, $\beta_0$ is the intercept, and $\beta_1, \ldots, \beta_{10}$ are the coefficients corresponding to the predictor variables.

>>>>

library(tidyverse)
library(lubridate)
library(tidyr)
library(dplyr)
library(tidymodels)
library(randomForest)
library(ranger)
library(caret)
library(themis)
library(yardstick)

@@

We find that danceability is the strongest predictor of whether or not a song has been produced by Antonoff. Furthermore, when the variance is considered along with other features of the dataset, the variance tends to be strongest classifier of whether a song was produced by Antonoff. Danceability in concert with other features in the dataset provides the best predictions.

@@

In total 3738 tracks from 828 different artists. We focus on identifying the styles of 7 producers: Jack Antonoff, Max Martin, Joel Little, Ariel Rechtshaid, Greg Kurstin, Paul Epworth and Rick Nowels. If tracks were produced by producers other than the main producers included in our study, they were marked as "other". We removed tracks from live performances, karaoke editions, international versions or remix albums. Where a deluxe album was available, we removed tracks belonging to the origin to avoid repetition.

@@

```{r}
#| label: fig-quadvalence
#| fig-cap: Valence and Energy Quadrant
#| echo: false
#| warning: false
av<- read_csv(here::here("inputs/data/antonoff_viz.csv"))

av$is_antonoff <- ifelse(av$producer == "antonoff", 1, 0)
av$is_antonoff <- as.factor(av$is_antonoff)

ggplot(av, aes(x = energy, y = valence)) +
  geom_point(aes(color = as.factor(is_antonoff)), alpha = 0.3) +
  geom_vline(xintercept = mean(df$energy), color = "gray") +
  geom_hline(yintercept = mean(df$valence), color = "gray") +
  theme_minimal()
```

@@

```{r}
#| label: fig-quaddanceability2
#| fig-cap: Danceability and Energy Quadrant
#| echo: false
#| warning: false

ggplot(av, aes(x = energy, y = danceability)) +
   geom_point(aes(color = as.factor(is_antonoff)), alpha = 0.3) +
  geom_vline(xintercept = mean(df$energy), color = "gray") +
  geom_hline(yintercept = mean(df$danceability),
              color = "gray") +
  theme_minimal()
```

@@

 av |>
   pivot_longer(
     cols = c(
       danceability,
       energy,
       loudness,
       mode,
       speechiness,
       acousticness,
       instrumentalness,
       liveness,
       valence,
       tempo
     ),
     names_to = "name",
     values_to = "value"
   ) |>
   ggplot(aes(x = value, color = is_antonoff)) +
   stat_ecdf() +
   facet_wrap(vars(name),
              nrow = 3,
              ncol = 4,
              scales = "free") +
   theme_minimal() +
   labs(x = "Value",
        y = "Proportion")

@@

av <- av |>
  mutate(is_antonoff= factor(is_antonoff))

av |>
  filter(album_release_date_precision == "day") |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(y = energy, x = album_release_date, color = is_antonoff)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080", "#FFB6C1")) +
  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Energy") +
  facet_wrap(~ artist_name, ncol = 3)
```

@@

av |>
  filter(album_release_date_precision == "day") |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(y = danceability, x = album_release_date, color = is_antonoff)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080", "#FFB6C1")) +
  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Danceability") +
  facet_wrap( ~ artist_name, ncol = 3)
```

@@

av |>
  filter(album_release_date_precision == "day") |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(y = valence, x = album_release_date, color = is_antonoff)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080", "#FFB6C1")) +
  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Valence") +
  facet_wrap( ~ artist_name, ncol = 3)
```

@@

av |>
  filter(album_release_date_precision == "day") |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(y = acousticness, x = album_release_date, color = is_antonoff)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080", "#FFB6C1")) +
  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Acousticness") +
  facet_wrap( ~ artist_name, ncol = 3)
```

@@

av |>
  filter(album_release_date_precision == "day") |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(y = instrumentalness, x = album_release_date, color = is_antonoff)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080", "#FFB6C1")) +
  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Instrumentalness") +
  facet_wrap( ~ artist_name, ncol = 3)
```

@@

av |>
  filter(album_release_date_precision == "day") |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(y = tempo, x = album_release_date, color = is_antonoff)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  scale_color_manual(values = c("#808080", "#FFB6C1")) +
  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Tempo") +
  facet_wrap( ~ artist_name, ncol = 3)
```

@@

$$
Pr(yi​=1)=logit−1(β0​+β1​×danceability+β2​×energy+...+β10​×tempo)
$$

where $\text{Pr}(y_i=1)$ is the probability of a song being produced by Antonoff, $\beta_0$ is the intercept, and $\beta_1, \ldots, \beta_{10}$ are the coefficients corresponding to the 10 predictor variables.
